---
title: "R users choropleth"
author: "Adolfo Álvarez"
date: "02/05/2015"
output: html_document
---

This markdown document includes the code necessary to generate the plot map from <a href="www.adolfoalvarez.cl", target="_blank">www.adolfoalvarez.cl</a>

Since there is one interesting map with the geolocalization of the groups in <a href="http://blog.revolutionanalytics.com/2015/04/revolution-analytics-microsoft.html", target="_blank">the blog of Revolution Analytics</a>, I will propose something different: Using <a href="https://github.com/hadley/rvest", target="_blank">rvest</a> for webscrapping, <a href="https://github.com/hadley/dplyr", target="_blank">dplyr</a> for data crunching and <a href="https://github.com/rstudio/leaflet", target="_blank">leaflet</a> for visualization, I want to create a choropleth map of R users per country, and obtain the top five groups by number of members. You can see similar choropleths maps generated in R <a href="http://rstudio-pubs-static.s3.amazonaws.com/58323_1911658f7b8841f1b14b349b3804540d.html", target="_blank">here</a> and <a href="http://rstudio-pubs-static.s3.amazonaws.com/52396_1455dccab1c3469a84171787db504502.html", target="_blank">here</a>

Let's get the data:

```{r, message=FALSE, warning=FALSE}
# library(devtools)
# install_github("hadley/rvest")
library(rvest)
library(dplyr)
r_users <- read_html("http://r-users-group.meetup.com/all/")
```

From the meetup directory I will extract the information of the groups using rvest and SelectorGadget as described in <a href="http://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html", target="_blank">this vignette</a>

```{r}
group <- r_users %>%
  html_nodes(".omngj_sj21") %>% #The name of the node is identified using "SelectorGadget"
  html_text()

latitude <- r_users %>%
  html_nodes(".latitude") %>%
  html_text()

longitude <- r_users %>%
  html_nodes(".longitude") %>%
  html_text()

country <- r_users %>%
  html_nodes(".country-name") %>%
  html_text()

locality <- r_users %>%
  html_nodes(".locality") %>%
  html_text()

members <- r_users %>%
  html_nodes(".note") %>%
  html_text()

data <- data.frame(group, locality, country, 
                   latitude, longitude, members, 
                   stringsAsFactors=F) %>%
  tbl_df()

data <- data %>%
  mutate(latitude=as.numeric(latitude),
         longitude=as.numeric(longitude),
         members=gsub(" members","",members) %>% as.numeric()           
  )
data %>%
  select(group, country, members)
```

Now for the choropleth map we use rgdal and leaflet.

```{r, fig.width=9.5, warning=FALSE, message=FALSE}
library(rgdal)
library(leaflet)

# We download the data for the map from naturalearthdata.com
url <- "http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip"
folder <- getwd() #set a folder where to download and extract the data
file <- basename(url) 
# download.file(url, file)
# unzip(file, exdir = folder)

#And read it with rgdal library
world <- readOGR(dsn = folder, 
                     layer = "ne_50m_admin_0_countries",
                     encoding = "latin1", #you may need to use a different encoding
                     verbose = FALSE)

#This function will extract the top 5 groups and return them as text to be included in the map
myfun <- function(x,y){
  a <- data.frame(x,y) %>%
    arrange(-y) %>%
    slice(1:5)
  return(paste0(a$x, ", ", a$y, " members", collapse="<br>"))
}

#Now we group by country and apply the function
data <- data %>%
  group_by(country) %>%
  summarise(top5=myfun(group, members),
            members=sum(members, na.rm=T)              
  ) %>%
  ungroup() %>%
  mutate(country=toupper(country))

#We need to reconvert to data.frame to merge it with the SpatialPolygons data frame "world"
data <- data.frame(data) 
world <- merge(world, data,
               by.x = "iso_a2",
               by.y = "country",
               sort = FALSE)

#Colour palette. Check ?colorQuantile and ?RColorBrewer for more options
pal <- colorQuantile("YlOrRd", NULL, n = 20)

#Pop up: The info displayed when click on a country
world_popup <- paste0("<strong>Country: </strong>", 
                        world$name, 
                        "<br><strong>", 
                        "Members", 
                        ": </strong>", 
                        world$members,
                        "<br><strong>", 
                        "Top 5 groups", 
                        ": </strong><br>",
                        world$top5
)

#Tiles coming from stamen.com
tiles <- "http://{s}.tile.stamen.com/toner-lite/{z}/{x}/{y}.png"
attribution <- 'Map tiles by <a href="http://stamen.com">Stamen Design</a>, under <a href="http://creativecommons.org/licenses/by/3.0">CC BY 3.0</a>. Map data by <a href="http://www.naturalearthdata.com/">Natural Earth</a>.'

#And finally the map:
leaflet(data = world) %>%
  addTiles(urlTemplate = tiles,  
           attribution = attribution) %>%
  setView(0, 0, zoom = 2) %>%
  addPolygons(fillColor = ~pal(world$members), 
              fillOpacity = 0.8, 
              color = "#000000", 
              weight = 1, 
              popup = world_popup)
```

